---
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FuzzyResampling

<!-- badges: start -->

[![R-CMD-check](https://github.com/mroman-ibs/FuzzyResampling/workflows/R-CMD-check/badge.svg)](https://github.com/mroman-ibs/FuzzyResampling/actions)
<!-- badges: end -->

The goal of FuzzyResampling, a library written in R, is to provide additional resampling procedures, apart from the classical bootstrap (i.e. Efron's approach, see [@Efron1994]), for fuzzy data.
In the classical approach, secondary samples are drawing with replacement from the initial sample.
Therefore most of these bootstrap samples contain repeated values. Moreover, if the  size of the primary sample is small then all secondary samples consist of only a few distinct values, which is a serious disadvantage.

To overcome this problem, special resampling algorithms for fuzzy data were introduced (see [@grzegorzewski2019; @grzegorzewskietal2020; 
@grzegorzewskiamcs2020;@Grzegorzewski2021; @romaniuk_hryniewicz]).
These methods randomly create values that are "similar" to values from the initial sample, but not exactly the same.
During the creation process, some of the characteristics of the initial fuzzy values are kept (e.g., the value, the width, etc.).
It was shown that these algorithms provide serious advantages in some statistical areas (like standard error
estimation or hypothesis testing) if they are compared with the classical approach.
For detailed information concerning the theoretical foundations and practical applications of these resampling methods please see the above-mentioned references.

The initial sample (in the form of a vector or a matrix) should consist of triangular or trapezoidal fuzzy numbers.

Some additional procedures related to these resampling methods are also provided, like calculation of the Bertoluzza et al.'s distance (aka the mid/spread distance, see [@Bertoluzza1995]), estimation of the p-value of the one-sample bootstrapped test for the mean (see [@lubiano2016]), and estimation of the standard error or the mean-squared error for the mean (see [@grzegorzewski2021]).

The following procedures are available in the library (R language):

+ Resampling procedures:
  + *classicalBootstrap* - classical approach based on Efron's method,
  + *VAmethod* - resampling method which preserves the value and ambiguity (see [@grzegorzewskietal2020]),
  + *EWmethod* - resampling method which preserves the expected value and width (see [@grzegorzewskiamcs2020]),
  + *VAAmethod* - resampling method which preserves the value, left-hand and right-hand ambiguities (see [@Grzegorzewski2021]),
  + *VAFmethod* - resampling method which preserves the value, ambiguity and fuzziness (see [@grzegorzewskietal2020]),
  + *dmethod* - resampling method which preserves the left end of the cores and increments (see [@romaniuk_hryniewicz]),
  + *wmethod* - resampling method which uses the special *w density* to "smooth" the output fuzzy value (see [@romaniuk_hryniewicz]).
  
+ Random generation of the initial samples:  
  + *GeneratorNU* - generation of the initial sample using the normal and uniform distributions (see [@grzegorzewskietal2020]),
  + *GeneratorNExpUU* - generation of the initial sample using the normal, exponential and uniform distributions (see [@grzegorzewskietal2020]).
  
+ Applications of the bootstrapped samples:  

  + *OneSampleCTest* - estimation of the p-value of the one-sample test for the mean (see [@lubiano2016]),
  + *SEResamplingMean* - estimation of the standard error or the mean-squared error for the mean (see [@grzegorzewski2021]).

+ Additional procedures:
  + *BertoluzzaDistance* - calculation of the Bertoluzza et al.'s distance (aka the mid/spread distance, see [@Bertoluzza1995]).

## Installation

You can install the latest development version of FuzzyResampling with:

``` r
library(devtools)
install_github("mroman-ibs/FuzzyResampling")
```

You can install the latest stable version from CRAN with:

``` r
install.packages("FuzzyResampling")
```

## Examples


``` {r}
# set seed

set.seed(12345)

# load library

library(FuzzyResampling)

# prepare some fuzzy numbers

fuzzyValues <- matrix(c(0.25,0.5,1,1.25,0.75,1,1.5,2.2,-1,0,0,2),ncol = 4,byrow = TRUE)

fuzzyValues

# seed PRNG

set.seed(12345)

# generate the secondary sample using the classical approach

classicalBootstrap(fuzzyValues)


# generate the secondary sample using the VA method

VAmethod(fuzzyValues)


# generate the secondary sample (6 fuzzy numbers) using the d-method

dmethod(fuzzyValues, b = 6)


# calculate the mid/spread distance between the first value
# (from the first row) and the second one (from the second row)

BertoluzzaDistance(fuzzyValues[1,],fuzzyValues[2,])


# seed PRNG

set.seed(1234)

# calculate the p-value using the classical (i.e. Efron's) bootstrap

OneSampleCTest(fuzzyValues, mu_0 = c(0,0.5,1,1.5))

# calculate the p-value using the VA resampling method

OneSampleCTest(fuzzyValues, mu_0 = c(0,0.5,1,1.5),resamplingMethod = VAmethod)


# seed PRNG

set.seed(1234)

# calculate the SE of the mean using the classical (i.e. Efron's) bootstrap

SEResamplingMean(fuzzyValues)

# calculate the SE of the mean using the VA resampling method

SEResamplingMean(fuzzyValues, resamplingMethod = VAmethod)

# calculate the MSE of the given mean using the classical (i.e. Efron's) bootstrap

SEResamplingMean(fuzzyValues, trueMean = c(0,0.5,1,2))

# calculate the MSE of the given mean using the VA resampling method

SEResamplingMean(fuzzyValues, resamplingMethod = VAmethod, trueMean = c(0,0.5,1,2))



```

``` r
# help concerning the VA method

?VAmethod

```

## References


