---
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FuzzyResampling

<!-- badges: start -->
<!-- badges: end -->

The goal of FuzzyResampling is to provide additional resampling procedures, apart from the classical bootstrap (i.e. Efron's
approach, see [@Efron1994]), for fuzzy data.
In the classical approach, secondary samples are drawing with replacement from the initial sample.
Therefore most of these bootstrap samples contain repeated values. Moreover, if the  size of the primary sample is small then all secondary samples consist of only a few distinct values, which is a serious disadvantage.

To overcome this problem, special resampling algorithms for fuzzy data were introduced (see [@grzegorzewski2019; @grzegorzewskietal2020; 
@grzegorzewskiamcs2020; @romaniuk_hryniewicz]).
These methods randomly create values that are "similar" to values from the initial sample, but not exactly the same.
During the creation process, some of the characteristics of the initial fuzzy values are kept (e.g., the value, the width, etc.).
It was shown that these algorithms provide serious advantages in some statistical areas (like standard error
estimation or hyphothesis testing) if they are compared with the classical approach.
For detailed information concerning the theoretical foundations and practical applications of these resampling methods please see the above-mentioned references.

The initial sample (in the form of a vector or a matrix) should consist of triangular or trapezoidal fuzzy numbers.

The following procedures are available in the library (R language):

+ *classicalBootstrap*
+ *VAmethod*
+ *EWmethod*
+ *VAAmethod*
+ *VAFmethod*
+ *dmethod*
+ *wmethod*

## Installation

You can install the released version of FuzzyResampling with:

``` r
library(devtools)
install_github("mroman-ibs/FuzzyResampling")
```

## Examples


``` {r}
# set seed

set.seed(12345)

# load library

library(FuzzyResampling)

# prepare some fuzzy numbers

fuzzyValues <- matrix(c(0.25,0.5,1,1.25,0.75,1,1.5,2.2,-1,0,0,2),ncol = 4,byrow = TRUE)

fuzzyValues

# generate the secondary sample using the classical approach

classicalBootstrap(fuzzyValues)

# generate the secondary sample using the VA method

VAmethod(fuzzyValues)


```

``` r
# help concerning the VA method

?VAmethod

```

## References


